---
title: "Week 02: COVID Visualization Activity"
date: "September, 11 2023"
author: "Hanwen Miao"
---

Today, we'll be working with a data set related to COVID.
This data is based on data from the the [COVID Tracking Project](https://covidtracking.com/).
I cleaned up this data and also added total populations from the 2020 for each of the relevant categories.
Note, due to differences in the way race and ethnicity are encoded in the census as compared to the the COVID Tracking Project, the population counts for LatinX may be somewhat inaccurate.

```{r load_data}
library(tidyverse)
load("CRDT Data.RData")
ls()
write.csv(covid_data_race, "covid_data_race.csv", row.names=FALSE)
```

I've include 4 different data sets.
They all have the same data but have it represented in different ways.
Try using the different data sets and see which ones are good for making which plots.


```{r first_plot}
ggplot(covid_data_count, aes(x = date, y = Cases)) + geom_point()
```



If you want to only look at a specific state, you can do it like this.
For now, see what you can do just using `ggplot`.

```{r}
covid_data_count |> 
  filter(state == "MA") |> 
  ggplot(aes(x = date, y = Cases, color = race)) + geom_line()
```

1. Write code for a plot and describe what you observe?
2. Are there any conclusions you can draw?
3. What were you not able to do due to not having the R knowledge?
4. What other data would be useful to better understand this data?
```{r}
# Load necessary libraries
library(tidyverse)

# Read the CSV file
covid_data_race <- read.csv("covid_data_race.csv")

# Filter for 'Cases' count type
cases_data <- subset(covid_data_race, count_type == "Cases")

# Summing counts by racial groups and removing NA values
racial_counts <- colSums(cases_data[,c("White_count", "Black_count", "Asian_count", 
                                       "AIAN_count", "NHPI_count", 
                                       "Multiracial_count", "Other_count")], na.rm = TRUE)

# Filter out non-finite values
racial_counts <- racial_counts[is.finite(racial_counts)]

# Rename the labels
names(racial_counts) <- c("White", "Black", "Asian", "AIAN", "NHPI", "Multiracial", "Other")

# Plot
racial_counts <- sort(racial_counts)
barplot(racial_counts, horiz = TRUE, las = 1, col = "blue",
        main = "Total COVID-19 Cases by Racial Group",
        xlab = "Total Cases",
        ylab = "Racial Group",
        border = "white",
        cex.names = 0.8)
```
Observations:
The White racial group has the highest number of reported COVID-19 cases, followed by the Black racial group. Asian, AIAN (American Indian/Alaska Native), and Multiracial groups have comparatively lower numbers of reported cases. The NHPI (Native Hawaiian and Other Pacific Islander) and Other racial groups have the least number of reported cases.
Conclusions:
There is a marked disparity in reported COVID-19 cases across different racial groups. The reasons for this disparity could be population distribution, social determinants of health, access to healthcare, and more. Addressing racial disparities in health outcomes would be essential. 
Limitation:
The data represents raw counts, and while population proportions have been provided, this analysis did not normalize the case counts by population. This normalization would provide a more accurate representation of case rates per capita across racial groups.
Socioeconomic data, such as income levels, housing density, and occupation types, could provide insights into why certain racial groups might be more exposed or vulnerable.

